\plaintitle

### 2025-05-21 docs/style/article-body.tex
pushing algorithm update

True random numbers, then, cannot be produced solely through code. These systems require some input that is neither replicable nor reproducible. One method that can be realistically used is the inherently random movement of lava lamps \cite{lavarand}, which is used as a backup source of randomness for Cloudflare\footnote{\href{https://blog.cloudflare.com/randomness-101-lavarand-in-production/}{Cloudflare.com, accessed 2025-03-10}}. Another proposed solution for this is quantum random number generation (\emph{QRNG}) \cite{QRNG}. By reading quantum fluctuation signals from any given source, for instance an optical signal, the inherent natural unpredictability of said signal can be harnessed in order to produce a random number from a state that is nigh impossible to reproduce accurately. Clason \cite{Clason2023} presents a device that generates a fluctuating analogue signal utilizing this method, which we will use as a basis for our work.

Further details about how this signal is produced and sampled will be introduced in section 2 and 3. The optical signal output by the device described in \cite{Clason2023} needs to be converted to a stream of random, raw bits via an Analog to Digital Converter (\emph{ADC}). Some post-processing of the said raw bits has to be performed in order to ensure that the bits are workable, and to remove potential deterministic patterns from the data. One method for this post-processing we will explore in this work is Toeplitz extraction \cite{toeplitz}, typically performed on the host computer utilizing the randomly generated numbers.

\[
    \text{OQRNG Analogue output} \rightarrow \text{ADC-converter} \rightarrow \text{post-processing}
\]

Clason proposes a simpler and cheaper way to achieve QRNG \cite{Clason2023}. In keeping with this, utilizing microcontrollers rather than a host computer for processing the raw bits extracted further helps to keep the costs low and the solution reasonably complex. Additionally, this allows the system to be self-contained, for instance a simple and portable USB-device which both generates and processes true random numbers. Further details regarding the microcontrollers used in our work will be outlined in section 3.

However, due to the lower processing power of the average microcontroller, any implementation of Toeplitz extraction needs to work quickly and efficiently despite this hardware constraint. Experimenting with efficient implementations of this well-defined algorithm is the main focus of this thesis, and the key research areas explored in this work are as follows:

@@

Toeplitz extraction been optimized quite well, and previous research can be utilized for this. However, there are still considerations when implementing the firmware for the microcontroller in order to optimize the code. Our goal is to attempt several implementations in order to find the most optimal implementation with the least amount of effective processing time spent on the algorithm.

@@

There will unequivocally be a bottleneck for the processing speed. For instance, the speed at which the ADC can process the optical signal into raw bits as well as the speed that the USB output can transfer processed random number to the host computer will be limiting factors. Further details on the limitations of the ADC will be outlined in section 3. The slowest of these bottlenecks will inevitably be the limiting factor for any implementation. Our research aims to ensure that our implementation of Toeplitz extraction does not become the limiting factor, but rather processing data fast enough to match or exceed the speed of the hardware.

@@

The idea of an optical QRNG (\emph{OQRNG}) is not a novel one. The basis of the theory is that intrinsically random properties of a quantum process. Stefanov et. al. \cite{StefanovOptical} proposes using the random choice of a photon between two output signals to generate a random stream of bits, however the theory behind it can be applied to other quantum processes as well. This particular theorem has been implemented by Wayne et. al. \cite{Wayne} to create a quantum number generator. While this article proves the efficacy of OQRNG, it utilizes a slightly different method.

@@

A more recent implementation of OQRNG in a smaller scale has been presented by Singh et. al. \cite{singh}. This particular implementation uses a bespoke circuit board where all components are present on a single board -- e.g., this experimental setup contains an integrated ADC, post-processor, entropy controller and entropy generator. While this article cements the viability of OQRNG using shot noise (despite the article not being confirmed as peer reviewed), the bespoke nature of the circuit board makes this experiment difficult to reproduce. As our thesis will use commercially available ADCs and microcontrollers, the only bespoke component is the shot noise generator itself. Furthermore, the Toeplitz extraction is not run on the microcontroller itself in these experiments -- instead, the hashing of these raw bits is done on the receiving computer as this bespoke circuit board featured a relatively weak processor.

In summary, previous research has proven that OQRNG can generate true randomness, and more specifically, Shen et. al. \cite{contender1} and Singh et. al. \cite{singh} both implement OQRNG through readings of shot noise. However, there are limitations in both of these works. Either the system that generates the shot noise is large and complex \cite{contender1} or the system is built on bespoke hardware with limitations in processing power which prevents a fully integrated system \cite{singh}. Furthermore, to the best of our knowledge, most of the work in this field is from the perspective of physicists, and there appears to be little research on this subject in the domain of computer science. Our work aims to bridge this gap by using commercially available hardware (other than the bespoke shot noise generator \cite{Clason2023}) and focuses on implementing Toeplitz extraction directly on the microcontroller. Rather than focusing on the intricacies of quantum fluctuations, we will instead approach this problem from a computer science perspective.

@@

Our work is a practical continuation of the work of Clason \cite{Clason2023}. In this work, quantum shot noise originating from photodiodes was studies, and in so doing a prototype device was constructed. This device read from an optical source, outputting analog voltage from the data ``seen'' by the diodes. A prototype was constructed, in which an LED is read by a photodiode soldered millimeters apart. In this section, we introduce the hardware used for our implementations as well as the considerations taken in order to shift the focus from physics to computer science.

@@

Whereas the exact quantum mechanisms that ensure that this system ensures randomness and further details regarding the OQRNG-device is better derived directly from Clasons work \cite{Clason2023}, the end result as it correlates to our study is an inherently random, analog voltage current.

@@

This analog current isn't suitable to operate on without further processing. As mentioned in section 1, the signal needs to pass through an ADC and converted into raw bits in order for it to be usable. In his thesis, Clason \cite{Clason2023} suggests a discrete ADC chip capable of analyzing frequencies higher than 25 MHz, as this is the highest frequency studied in his work. However, in the interest of keeping the implementation light and cheap, we will be using ADCs that provide less samples per second and lower frequencies. This is done mainly for ease of development and access to this hardware.

Many microcontrollers furthermore come equipped with internal ADCs that can be utilized, and while these provide a lower sample size (often around 1 MSPS), the ease of development may be prudent to utilize for this proof-of-concept. While our initial ADC has a fairly low throughput, this can always be upgraded if it ends up becoming too limiting.

Should these internal ADCs prove too limiting, we propose utilizing MAX11102AUB\footnote{\href{https://www.farnell.com/datasheets/1913106.pdf}{Technical specification for MAX11102AUB, accessed 2025-03-13}} with an effective sample rate of 2 million samples per second (MSPS). This ADC provides a 12 bit sample size, providing roughly 24 Mbit/s of sampled data per second, derived by the following calculation.

@@

Microcontrollers (MCUs) are compact and low-power computing devices designed for embedded systems and real-time operations, and suitable as a processing unit for the purposes of this work. Unlike general purpose CPUs, an MCU integrates a processor, memory and peripherals (\emph{such as an ADC, for instance}) into a single chip. Furthermore, modern MCUs often feature advanced microarchitectural elements to enhance processing capabilities on single threads (\emph{such as dual-issue superscalar architectures, allowing the MCU to run several instructions per CPU cycle}), making them suitable candidates for the post-processing required for OQRNG-data.

Since MCUs often function under strict timing requirements, it is critical to have effective ways to access memory and transfer data for processing in real time. High-performance MCUs enhance memory usage in various ways. Some of them use Tightly Coupled Memory\footnote{\href{https://www.farnell.com/datasheets/1913106.pdf}{Technical specification for MAX11102AUB, accessed 2025-03-13}} (TCM), which gives fast SRAM with specific access routes for important data, avoiding cache misses and guaranteeing consistent performance. Moreover, instruction and data caching techniques, including instruction prefetching and branch prediction, help minimize execution delays in computationally intensive real-time applications. Another important aspect is Direct Memory Access\footnote{\href{https://www.sciencedirect.com/topics/computer-science/direct-memory-access}{ScienceDirect Journals \& Books, accessed 2025-03-13}} (DMA), which enables data transfer between peripherals such as the ADC and RAM, without CPU intervention. This offloading reduces processing overhead, allowing the MCU to manage fast data transfers effectively. These improvements are especially significant for Toeplitz extraction, where large amount of random data needs to be processed and sent quickly with low delay. Efficient memory management guarantees that randomness extraction can occur rapidly without major slowdowns in computing. Both approaches will be tested during development.

In our work, we intend to use Teensy 4.1\footnote{\href{https://www.pjrc.com/store/teensy41.html}{Teensy developer documentation, accessed 2025-02-27}}, based on the ARM Cortex-M7. This MCU is especially suitable for computationally demanding tasks involving randomness extraction due to its dual-issue superscalar architecture and Digital Signal Processing (\emph{DSP}) capabilities. The floating-point unit\footnote{\href{https://www.sciencedirect.com/topics/computer-science/floating-point-unit}{ScienceDirect Journals \& Books, accessed 2025-03-13}} (FPU) and Single Instruction, Multiple Data (\emph{SIMD}) style DSP instructions improves how quickly it can perform bitwise and arithmetic tasks, which are crucial for quick Toeplitz extraction. SIMD-controlled DSP architectures, as described by Han et al. \cite{simd-dsp}, leverage parallel vectorized computation to accelerate matrix operations -- making them highly effective for Toeplitz matrix-vector multiplications.

@@

The raw bits from the ADC can potentially have some deterministic patterns, and as such have to be processed somehow in order to remove these patterns. Several methods exist for this purpose, and for our work, we will perform this preprocessing via Toeplitz extraction. The main focus of this study is to implement this extraction algorithm as effectively as possible on resource constrained hardware.

@@

To summarize the theoretical working of Toeplitz extraction (as explained by Chouhan et al. \cite{toeplitz-desc}), a pre-determined seed matrix (\(T\)) is multiplied with the sampled raw bit matrix (\(K\)). The size of the seed is directly dependent on the size of the sampled data, and can be fixed or continually re-sampled as needed. To ensure high levels of entropy, our intuition is that re-sampling the seed from the OQRNG-device continually is prudent. The sample and seed with then be processed with matrix multiplication to remove deterministic patterns, and produce a bitstring that results in our randomly generated number. An example of how this extraction works can be seen below.

\begin{verbatim}
Where n = number of input bits,
      m = number of output bits:

Let X be the input array (length n)
Let T be the seed matrix (length n + m – 1)
Let Y be a new array of length m

For i from 0 to m - 1:
  sum = 0
  For j from 0 to n - 1:
    sum = sum + X[j] * T[i + j]
  Y[i] = sum mod 2

Return Y
\end{verbatim}

@@

With the assumption that the OQRNG-device produces a truly random analog signal, we can now clearly define the scope in which this thesis operates. Considering the maximum conversion speed from the ADC and the USB-output from the MCU, we have a clear bound over 24 Mbit/s (imposed by the ADC) in which Toeplitz extraction needs to be executed. Any speeds over 2.86 MB/s allows us to upgrade the ADC iteratively to continue increasing the output speed. Any implementation of Toeplitz extraction must then execute fast enough on any given microcontroller feasible for the proposed quantum RNG-thumbstick as to not be the decisive limiting factor.

@@

Another important point that is often neglected in randomness extraction using Toeplitz matrices is how the seed key is handled when forming the Toeplitz matrix. Numerous systems, such as those developed by Chouhan et al. \cite{toeplitz-desc} and Zhang et al. \cite{toeplitz}, depends on fixed seeds. Fixed seeds can create security risks over prolonged operation time. To tackle this problem, Lin et al \cite{lin} proposed a method for seed-renewable Toeplitz post-processing in QRNG. Their strategy incorporates a dynamic seed pool within the FPGA, where each instance of post-processing picks a new, randomly selected seed. Thus, minimizing temporal correlations between extractions. Furthermore, an external seed updating mechanism via PCIe ensures that seeds are refreshed whenever a certain security limit is reached. Compared to fixed-seed methods, this renewable approach enhances cryptographic robustness and ensures sustained high-security randomness extraction in real-world applications.

Efficient Toeplitz matrix-vector multiplication (TMVM) is critical for optimizing randomness extraction that relies on Toeplitz, particulary in constrained hardware environments. Liao et al. \cite{liao} showed that this process could be greatly accelerated using Fast Fourier Transform (FFT) and its inverse (IFFT). Thus, reducing computational complexity from \(O(n^2)\) to \(O(n log n)\). Their implementation on FPGA utilized this approach for deep neural networks, resulting in a 28.7 times decrease in model size while still achieving fast inference speeds. By using FFT and IFFT acceleration, Toeplitz post-processing for randomness extraction could achieve higher throughput. Thus, potentially could improve performance and reduce latency.

@@

Hyncica et. al. \cite{micromeasurements} propose that measuring execution time of algorithms directly via the microcontrollers internal timers (while subtracting the interrupt overhead) provides adequate measurements of an algorithms execution speed. An additional advantage is that the same code can be used to measure execution speed on several different microcontrollers, rather than relying on counting CPU cycles (as the process for this may vary greatly between controllers). As we will use fixed-size bitstrings for evaluation, we can then derive the throughput of the algorithm in \(Mbit/s\) as follows:

\[
\text{Throughput}{\mathrm{Mbit/s}}
= \frac{DataSize{\mathrm{bits}}}{ExecutionTime_{\mathrm{ms}}}
\times 10^{-3}
\phantom{12}(2)
\]

@@

\[
T_{\mathrm{ms}}
= \frac{64}{24}\times10^{-3}\ \mathrm{ms}
\approx 2.667\times10^{-3}\ \mathrm{ms}
=2.667\ \mu\mathrm{s}.
\phantom{12}(3)
\]

@@

\subsubsection{5.3 Limitations}\label{limitations}

Our proposed iterations all assume that the limited hardware will support it. Whereas we are confident that Teensy 4.1 will be able to handle each iteration step (even the naive implementation), the remaining microcontrollers with lower specifications might not be suitable for the first iterations. Testing the implementations on different microcontrollers could turn out to be unfeasible -- however, this remains to be seen during the experimentation.

@@

Table 1 presents the execution speeds of each iteration as listed in sections 5.3 and 5.4, alongside the exectution speed in Mbit/s as calculated by utilizing the average value in (3). My thought then is that we here talk about the results and what they demonstrate, before moving on to draw conclusions and attempt to answer our research questions in section 7.

>>>>

True random numbers, then, cannot be produced solely through code. These systems require some input that is neither replicable nor reproducible. One method that can be realistically used is the inherently random movement of lava lamps \cite{lavarand}, which is used as a backup source of randomness for Cloudflare\footnote{\href{https://blog.cloudflare.com/randomness-101-lavarand-in-production/}{Cloudflare.com, accessed 2025-03-10}}. Another proposed solution for this is quantum random number generation (\emph{QRNG}) \cite{QRNG}. By reading quantum fluctuation signals from any given source, for instance an optical signal, the inherent natural unpredictability of said signal can be harnessed in order to produce a random number from a state that is nigh impossible to reproduce accurately. Clason \cite{Clason2023} presents a device that generates a fluctuating analogue signal utilizing this method.

The optical signal output by the device described in \cite{Clason2023} needs to be converted to a stream of random, raw bits via an Analog to Digital Converter (\emph{ADC}). Some post-processing of the said raw bits has to be performed in order to ensure that the bits are workable, and to remove potential deterministic patterns from the data. One method for this post-processing we will explore in this work is Toeplitz extraction \cite{toeplitz}, typically performed on the host computer utilizing the randomly generated numbers. The post-processing finally yields a random number. This relationship can be seen in Figure 1.

\begin{figure}[ht] \centering \begin{tikzpicture}[>=latex] % top row, 2 cm apart
\node (A) {RNG-module}; \node (B) [right=2cm of A] {ADC-conversion};

    \node (C) [below=1cm of B] {Post-processing};
    \node (D) [left=2cm of C]  {Random number};

    \draw[->] (A) -- (B);
    \draw[->] (B) -- (C);
    \draw[->] (C) -- (D);

\end{tikzpicture} \caption{An abstract of the steps required for OQRNG.}
\label{fig:linear-system} \end{figure}

Clason proposes a simpler and cheaper way to achieve QRNG \cite{Clason2023} than what has been done in previous research. These systems have generally been large, bulky and expensive, whereas the method proposed is portable enough to theoretically be installed in a portable device -- for instance a USB-thumbstick. In keeping with this, utilizing microcontrollers rather than a host computer for processing the raw bits extracted allows this system to be self-contained and portable, and further helps to keep the costs low and the solution reasonably complex.

However, due to the limited processing power of the average microcontroller, any implementation of Toeplitz extraction needs to work quickly and efficiently despite this hardware constraint. Experimenting with efficient implementations of this well-defined algorithm is the main focus of this thesis, and the key research areas explored in this work are as follows:

@@

Toeplitz extraction has been optimized quite well, and previous research can be utilized to address this research question. However, there are still considerations when implementing the firmware for the microcontroller in order to optimize the code. Our goal is to attempt several implementations in order to find the most optimal implementation with the least amount of effective processing time spent on the algorithm.

@@

There will unequivocally be a bottleneck for the processing speed. For instance, the speed at which the ADC can process the optical signal into raw bits as well as the speed that the USB output can transfer processed random number to the host computer will be limiting factors. Further details on the limitations of the ADC will be outlined in Section 3. The slowest of these bottlenecks will inevitably be the limiting factor for any implementation. Our research aims to ensure that our implementation of Toeplitz extraction does not become the limiting factor, but rather processing data fast enough to match or exceed the speed of the hardware.

@@

The idea of an optical QRNG (\emph{OQRNG}) is not a novel one. The basis of the theory is the intrinsically random properties of a quantum process. Stefanov et. al. \cite{StefanovOptical} proposes using the random choice of a photon between two output signals to generate a random stream of bits, however the theory behind it can be applied to other quantum processes as well. This particular theorem has been implemented by Wayne et. al. \cite{Wayne} to create a quantum number generator. While this article proves the efficacy of OQRNG, it utilizes a slightly different method.

@@

A more recent implementation of OQRNG in a smaller scale has been presented by Singh et. al. \cite{singh}. This particular implementation uses a bespoke circuit board where all components are present on a single board -- e.g., this experimental setup contains an integrated ADC, post-processor, entropy controller and entropy generator. While this article cements the viability of OQRNG using shot noise (\emph{despite the article not being confirmed as peer reviewed}), the bespoke nature of the circuit board makes this experiment difficult to reproduce. As our thesis will use commercially available ADCs and microcontrollers, the only bespoke component is the shot noise generator itself. Furthermore, the Toeplitz extraction is not run on the microcontroller itself in these experiments -- instead, the hashing of these raw bits is done on the receiving computer as this bespoke circuit board featured a relatively weak processor.

In summary, previous research has proven that OQRNG can generate true randomness, and more specifically, Shen et. al. \cite{contender1} and Singh et. al. \cite{singh} both implement OQRNG through readings of shot noise. However, there are limitations in both of these works. Either the system that generates the shot noise is large and complex \cite{contender1} or the system is built on bespoke hardware with limitations in processing power which prevents a fully integrated system \cite{singh}. Furthermore, to the best of our knowledge, most of the work in this field is from the perspective of physicists, and there appears to be little research on this subject in the domain of computer science. Our work aims to bridge this gap by using commercially available hardware (\emph{other than the bespoke shot noise generator \cite{Clason2023}}) and focuses on implementing Toeplitz extraction directly on the microcontroller. Rather than focusing on the intricacies of quantum fluctuations, we will instead approach this problem from a computer science perspective.

@@

Our work is a practical continuation of the work of Clason \cite{Clason2023}. In this work, quantum shot noise originating from photodiodes was studied, and in so doing a prototype device was constructed. This prototype utilized an LED soldered millimeters apart from a photodiode, which in turn outputs a variable analog signal from the device. In this section, we introduce the remaining hardware used for our implementations as well as the considerations taken in order to shift the focus from physics to computer science.

@@

Whereas the exact quantum mechanisms that ensure that this system ensures randomness and further details regarding the OQRNG-device is better derived directly from Clasons work \cite{Clason2023}, the primary concern for our study is the inherently random, analog voltage current produced by the system.

@@

This analog current is not suitable to operate on without further processing. As mentioned in Section 1, the signal needs to pass through an ADC to be converted into raw bits. In his thesis, Clason \cite{Clason2023} suggests a discrete ADC chip capable of analyzing frequencies higher than 25 MHz, as this is the highest frequency studied in his work. However, in the interest of keeping the implementation light and cheap, we will be using ADCs that provide less samples per second and lower frequencies. This is done mainly for ease of development and access to this hardware.

Many microcontrollers furthermore come equipped with internal ADCs that can be utilized, and while these provide a lower sample size (\emph{often around 1 MSPS}), the ease of development may be prudent to utilize for this proof-of-concept. While our initial ADC has a fairly low throughput, this can always be upgraded if it ends up becoming too limiting.

Should these internal ADCs prove too limiting, we propose utilizing MAX11102AUB\footnote{\href{https://www.farnell.com/datasheets/1913106.pdf}{Technical specification for MAX11102AUB, accessed 2025-03-13}} with an effective sample rate of 2 million samples per second (\emph{MSPS}). This ADC provides a 12 bit sample size, providing roughly 24 Mbit/s of sampled data per second, derived by the following calculation.

@@

Microcontrollers (\emph{MCUs}) are compact and low-power computing devices designed for embedded systems and real-time operations, and suitable as a processing unit for the purposes of this work. Unlike general CPUs, an MCU integrates a processor, memory and peripherals (\emph{such as an ADC}) into a single chip. Furthermore, modern MCUs often feature advanced microarchitectural elements to enhance processing capabilities on single threads (\emph{such as dual-issue superscalar architectures, allowing the MCU to run several instructions per CPU cycle}), making them suitable candidates for the post-processing required for OQRNG-data.

Since MCUs often function under strict timing requirements, it is critical to have effective ways to access memory and transfer data for processing in real time. High-performance MCUs enhance memory usage in various ways. Some of them use Tightly Coupled Memory\footnote{\href{https://www.farnell.com/datasheets/1913106.pdf}{Technical specification for MAX11102AUB, accessed 2025-03-13}} (\emph{TCM}), which gives fast SRAM with specific access routes for important data, avoiding cache misses and guaranteeing consistent performance. Moreover, instruction and data caching techniques, including instruction pre-fetching and branch prediction, help minimize execution delays in computationally intensive real-time applications. Another important aspect is Direct Memory Access\footnote{\href{https://www.sciencedirect.com/topics/computer-science/direct-memory-access}{ScienceDirect Journals \& Books, accessed 2025-03-13}} (DMA), which enables data transfer between peripherals such as the ADC and RAM, without CPU intervention. This offloading reduces processing overhead, allowing the MCU to manage fast data transfers effectively. These improvements are especially significant for Toeplitz extraction, where large amount of random data needs to be processed and sent quickly with low delays. Efficient memory management guarantees that randomness extraction can occur rapidly without major slowdowns in computing. Both approaches will be tested during development.

In our work, we intend to use Teensy 4.1\footnote{\href{https://www.pjrc.com/store/teensy41.html}{Teensy developer documentation, accessed 2025-02-27}}, based on the ARM Cortex-M7. This MCU is especially suitable for computationally demanding tasks involving randomness extraction due to its dual-issue superscalar architecture and Digital Signal Processing (\emph{DSP}) capabilities. The floating-point unit\footnote{\href{https://www.sciencedirect.com/topics/computer-science/floating-point-unit}{ScienceDirect Journals \& Books, accessed 2025-03-13}} (\emph{FPU}) and Single Instruction, Multiple Data (\emph{SIMD}) style DSP instructions improve how quickly it can perform bitwise and arithmetic tasks, which is crucial for quick Toeplitz extraction. SIMD-controlled DSP architectures, as described by Han et al. \cite{simd-dsp}, leverage parallel vectorized computation to accelerate matrix operations -- making them highly effective for Toeplitz matrix-vector multiplications.

@@

The raw bits from the ADC can potentially have some deterministic patterns, and as such have to be processed somehow in order to remove these patterns. Several methods exist for this purpose, and for our work, we will perform this pre-processing via Toeplitz extraction. The main focus of this study is to implement this extraction algorithm as effectively as possible on resource constrained hardware.

@@

To summarize the theoretical working of Toeplitz extraction (\emph{as explained by Chouhan et al. \cite{toeplitz-desc}}), a pre-determined seed matrix (\(T\)) is multiplied with the sampled raw bit matrix (\(K\)). The size of the seed is directly dependent on the size of the sampled data, and can be fixed or continually re-sampled as needed. To ensure high levels of entropy, our intuition is that re-sampling the seed from the OQRNG-device continually is prudent. The sample and seed will then be processed with matrix multiplication to remove deterministic patterns, and produce a bitstring that results in our randomly generated number. An example of how this extraction works can be seen below.

\begin{algorithm}[ht] \caption{Toeplitz extraction}\label{alg:bit-conv}
\begin{algorithmic}[1] \REQUIRE \(x[0\,..\,n-1]\) \Comment{input bit array of
length \(n\)} \REQUIRE \(t[0\,..\,n+m-2]\) \Comment{seed matrix of length
\(n+m-1\)} \ENSURE \(y[0\,..\,m-1]\) \Comment{output bit array of length \(m\)}
\FOR{\(i = 0\) to \(m - 1\)} \STATE sum = 0 \FOR{\(j = 0\) to \(n - 1\)} \STATE
sum = sum + \(x[j] \* t[i + j]\) \ENDFOR \STATE \(y[i] =
\mathrm{sum}\;\bmod\;2\) \ENDFOR \RETURN \(y\) \end{algorithmic} \end{algorithm}

@@

With the assumption that the OQRNG-device produces a truly random analog signal, we can now clearly define the scope in which this thesis operates. Considering the maximum conversion speed from the ADC and the USB-output from the MCU, we have a clear bound over 24 Mbit/s (\emph{imposed by the ADC}) in which Toeplitz extraction needs to be executed. Any speeds over 2.86 MB/s allows us to upgrade the ADC iteratively to continue increasing the output speed. Any implementation of Toeplitz extraction must then execute fast enough on any given microcontroller feasible for the proposed quantum RNG-thumbstick as to not be the decisive limiting factor.

@@

Another important point that is often neglected in randomness extraction using Toeplitz matrices is how the seed key is handled when forming the Toeplitz matrix. Numerous systems, such as those developed by Chouhan et al. \cite{toeplitz-desc} and Zhang et al. \cite{toeplitz}, depends on fixed seeds. Fixed seeds can create security risks over prolonged operation time. To tackle this problem, Lin et al \cite{lin} proposed a method for seed-renewable Toeplitz post-processing in QRNG. Their strategy incorporates a dynamic seed pool within the FPGA, where each instance of post-processing picks a new, randomly selected seed. Thus, minimizing temporal correlations between extractions. Furthermore, an external seed updating mechanism via Peripheral Component Interconnect Express (\emph{PCIe}) ensures that seeds are refreshed whenever a certain security limit is reached. Compared to fixed-seed methods, this renewable approach enhances cryptographic robustness and ensures sustained high-security randomness extraction in real-world applications.

Efficient Toeplitz matrix-vector multiplication (\emph{TMVM}) is critical for optimizing randomness extraction that relies on Toeplitz, particulary in constrained hardware environments. Liao et al. \cite{liao} showed that this process could be greatly accelerated using Fast Fourier Transform (\emph{FFT}) and its inverse (\emph{IFFT}) -- reducing computational complexity from \(O(n^2)\) to \(O(n \log n)\). Their implementation on FPGA utilized this approach for deep neural networks, resulting in a 28.7 times decrease in model size while still achieving fast inference speeds. By using FFT and IFFT acceleration, Toeplitz post-processing for randomness extraction could achieve higher throughput. Thus, potentially could improve performance and reduce latency.

@@

Hyncica et. al. \cite{micromeasurements} propose that measuring execution time of algorithms directly via the microcontrollers internal timers (while subtracting the interrupt overhead) provides adequate measurements of the execution speed of an algorithm. An additional advantage is that the same code can be used to measure execution speed on several different microcontrollers, rather than relying on counting CPU cycles (as the process for this may vary greatly between controllers). As we will use fixed-size bitstrings for evaluation, we can then derive the throughput of the algorithm in \(Mbit/s\) as follows:

\[
Throughput_\mathrm{Mbit/s}
= \frac{DataSize_\mathrm{bits}}{ExecutionTime_\mathrm{ms}}
\times 10^{-3}
\phantom{12}(2)
\]

@@

\[
\frac{64}{24}\times10^{-3}\ \mathrm{ms}
\approx 2.667\times10^{-3}\ \mathrm{ms}
=2.667\ \mu\mathrm{s}.
\phantom{12}(3)
\]

@@

\subsubsection{5.5 Limitations}\label{limitations}

Our proposed iterations all assume that the limited hardware will support it. Whereas we are confident that Teensy 4.1 will be able to handle each iteration

step (\emph{even the naive implementation}), the remaining microcontrollers with lower specifications might not be suitable for the first iterations. Testing the implementations on different microcontrollers could turn out to be unfeasible -- however, this remains to be seen during the experimentation.

@@

Table 1 presents the execution speeds of each iteration as listed in sections 5.3 and 5.4, alongside the execution speed in Mbit/s as calculated by utilizing the average value in (3). My thought then is that we here talk about the results and what they demonstrate, before moving on to draw conclusions and attempt to answer our research questions in section 7.
